---
title: "Replicating Caliendo and Parro 2015"
output:
  litedown::html_format:
    options:
      toc: true
      number_sections: true
vignette: >
  %\VignetteIndexEntry{Example: replicating NAFTA tariff reductions}
  %\VignetteEngine{litedown::vignette}
  %\VignetteEncoding{UTF-8}
editor:
  markdown:
    wrap: sentence
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Trade and welfare effects from NAFTA's tariff reductions (section 5.1)

This vignette uses `cp2015` to reproduce the simulation results presented in
section 5.1 of Caliendo and Parro (2015), "Estimates of the Trade and Welfare
Effects of NAFTA".

```{r setup}
library(cp2015)
library(data.table)

options(knitr.kable.NA = "")
```

## Data

The package includes `cp2015nafta`, a list containing the required model sets,
intermediate consumption, final consumption, value added, bilateral trade and
tariff data, deficits, and trade elasticities.

```{r}
str(cp2015nafta)
```

## Welfare effects from NAFTA's tariff reductions (table 2)

Caliendo and Parro (2015) impose zero aggregate deficits in the baseline and
counterfactual scenarios. Use `zero_aggregate_deficit = TRUE` to reproduce that
scenario.

```{r}
results <- run_cp2015(data = cp2015nafta, zero_aggregate_deficit = TRUE)
str(results)
```

The welfare decomposition includes terms of trade, volume of trade, and
technical efficiency. The latter is zero in this tariff-only example.

```{r}
nafta <- c("Canada", "Mexico", "USA")
welfare <- results$welfare[region %in% nafta]

knitr::kable(
  welfare,
  digits = 2,
  col.names = c(
    "Country", "Terms of trade", "Volume of trade", "Technical efficiency",
    "Total", "Real wages"
  ),
  caption = "Welfare effects from NAFTA's tariff reductions"
)
```

## Bilateral welfare effects from NAFTA's tariff reductions (table 3)

```{r}
tot_bilateral <- results$tot[region %in% nafta]
tot_bilateral[, partner := fifelse(partner %in% nafta, "NAFTA", "ROW")]
tot_bilateral <- tot_bilateral[, .(ToT = sum(tot)), by = .(region, partner)]

vot_bilateral <- results$vot[region %in% nafta]
vot_bilateral[, partner := fifelse(partner %in% nafta, "NAFTA", "ROW")]
vot_bilateral <- vot_bilateral[, .(VoT = sum(vot)), by = .(region, partner)]

bilateral <- merge(
  tot_bilateral,
  vot_bilateral,
  by = c("region", "partner"),
  all = TRUE,
  sort = FALSE
)

bilateral <- dcast(
  bilateral,
  region ~ partner,
  value.var = c("ToT", "VoT")
)

knitr::kable(
  bilateral,
  digits = 2,
  caption = "Bilateral welfare effects from NAFTA's tariff reductions"
)
```

## Sectoral contribution to welfare effects from NAFTA's tariff reductions (table 4)

```{r}
tot_sector <- results$tot[region %in% nafta]
tot_sector <- tot_sector[, .(ToT = sum(tot)), by = .(region, sector)]
tot_sector[, ToT := 100 * ToT / sum(ToT), by = region]

vot_sector <- results$vot[region %in% nafta]
vot_sector <- vot_sector[, .(VoT = sum(vot)), by = .(region, sector)]
vot_sector[, VoT := 100 * VoT / sum(VoT), by = region]

sectoral <- merge(
  tot_sector,
  vot_sector,
  by = c("region", "sector"),
  all = TRUE,
  sort = FALSE
)

sectoral <- dcast(
  sectoral,
  sector ~ region,
  value.var = c("ToT", "VoT")
)

setcolorder(
  sectoral,
  c(
    "sector",
    "ToT_Mexico", "VoT_Mexico",
    "ToT_Canada", "VoT_Canada",
    "ToT_USA", "VoT_USA"
  )
)

knitr::kable(
  sectoral,
  digits = 2,
  col.names = c(
    "Sector",
    "Mexico: Terms of trade", "Mexico: Volume of trade",
    "Canada: Terms of trade", "Canada: Volume of trade",
    "United States: Terms of trade", "United States: Volume of trade"
  ),
  caption = "Sectoral contribution to welfare effects from NAFTA's tariff reductions"
)
```

## Trade effects from NAFTA's tariff reductions (table 5)

```{r}
trade <- results$trade
trade <- trade[importer %in% nafta & exporter %in% nafta]
trade <- trade[, .(trade_bln = sum(trade_bln), trade_cfl = sum(trade_cfl)),
  by = .(importer, exporter)
]
trade[, change := (trade_cfl / trade_bln - 1) * 100]
trade[exporter == importer, change := NA_real_]

knitr::kable(
  trade[, .(importer, exporter, change)],
  digits = 2,
  caption = "Trade effects from NAFTA's tariff reductions"
)
```
