---
title: "Replicating 'Estimates of the Trade and Welfare Effects of NAFTA Get access Arrow' by Caliendo and Parro (2015)"
author: "Mauricio 'Pacha' Vargas Sepulveda"
date: "2026-09-22"
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
bibliography: bibliography.bib
---

# Introduction

@caliendo2015estimates evaluate the economic impacts of the North American Free Trade Agreement using a multi-sector
Ricardian trade model incorporating intermediate goods and input-output linkages. Their findings show that Mexico's
welfare increases by 1.31%, USA's welfare increases by 0.08%, and Canada's welfare declines by 0.06% while at the same
time intra-bloc trade increases by 118% for Mexico, 41% for the USA and 11% for Canada.

This replication effort uses @vargas2025armadillo4r to leverage Armadillo's cube class (quasi third-order tensor, or
3D-matrix), which simplifies solving non-linear systems as the ones required for the Ricardian model with trade sectors.

The tables below correspond to an independent simulation of the results from section 5.1. in @caliendo2015estimates.

```{r}
#| warning: false
#| message: false
library(cp2015)
library(data.table)
options(knitr.kable.NA = "")
```

The `cp2015` package (\url{https://github.com/pachadotdev/cp2015/blob/main/README.md}) includes `cp2015nafta`, a list
containing the required model sets, intermediate consumption, final consumption, value added, bilateral trade and
tariff data, deficits, and trade elasticities.

For different datasets, such as those with trade figures from the OECD, it is required to reshape the dataset before
passing it to the package functions as in the structure below:

```{r}
str(cp2015nafta)
```

@caliendo2015estimates impose zero aggregate deficits in the baseline and counterfactual scenarios. We use
`zero_aggregate_deficit = TRUE` to reproduce that scenario.

```{r}
results <- run_cp2015(data = cp2015nafta, zero_aggregate_deficit = TRUE)
```

The welfare decomposition includes terms of trade, volume of trade, and technical efficiency. The latter is zero in this
tariff-only example. From the model results, we need to aggregate its outputs in different ways to replicate tables
two to five.

# Welfare effects from NAFTA's tariff reductions (Table 2)

```{r}
nafta <- c("Mexico", "Canada", "USA")
welfare <- results$welfare[region %in% nafta]
welfare[, tech := NULL]
welfare <- welfare[match(nafta, region), ]

knitr::kable(
  welfare,
  digits = 2,
  latex = TRUE,
  col.names = c("Country", "Total", "Terms of trade", "Volume of trade", "Real wages"),
  caption = "Welfare effects from NAFTA's tariff reductions (Table 2)"
)
```

# Bilateral welfare effects from NAFTA's tariff reductions (Table 3)

```{r}
tot_bilateral <- results$tot[region %in% nafta]
tot_bilateral[, partner := fifelse(partner %in% nafta, "NAFTA", "ROW")]
tot_bilateral <- tot_bilateral[, .(ToT = sum(tot)), by = .(region, partner)]

vot_bilateral <- results$vot[region %in% nafta]
vot_bilateral[, partner := fifelse(partner %in% nafta, "NAFTA", "ROW")]
vot_bilateral <- vot_bilateral[, .(VoT = sum(vot)), by = .(region, partner)]

bilateral <- merge(
  tot_bilateral,
  vot_bilateral,
  by = c("region", "partner"),
  all = TRUE,
  sort = FALSE
)

bilateral <- dcast(
  bilateral,
  region ~ partner,
  value.var = c("ToT", "VoT")
)

bilateral <- bilateral[match(nafta, region), ]

knitr::kable(
  bilateral,
  digits = 2,
  latex = TRUE,
  col.names = c("Country", "ToT NAFTA", "ToT ROW", "VoT NAFTA", "VoT ROW"),
  caption = "Bilateral welfare effects from NAFTA's tariff reductions
    (Table 3)"
)
```

# Sectoral contribution to welfare effects from NAFTA's tariff reductions (Table 4)

```{r}
tot_sector <- results$tot[region %in% nafta]
tot_sector <- tot_sector[, .(ToT = sum(tot)), by = .(region, sector)]
tot_sector[, ToT := 100 * ToT / sum(ToT), by = region]

vot_sector <- results$vot[region %in% nafta]
vot_sector <- vot_sector[, .(VoT = sum(vot)), by = .(region, sector)]
vot_sector[, VoT := 100 * VoT / sum(VoT), by = region]

sectoral <- merge(
  tot_sector,
  vot_sector,
  by = c("region", "sector"),
  all = TRUE,
  sort = FALSE
)

sectoral <- dcast(
  sectoral,
  sector ~ region,
  value.var = c("ToT", "VoT")
)

setcolorder(
  sectoral,
  c("sector", "ToT_Mexico", "VoT_Mexico", "ToT_Canada", "VoT_Canada",
    "ToT_USA", "VoT_USA")
)

knitr::kable(
  sectoral,
  digits = 2,
  latex = TRUE,
  col.names = c("Sector", gsub("_", " ", colnames(sectoral)[-1])),
  caption = "Sectoral contribution to welfare effects from NAFTA's
    tariff reductions (Table 4)"
)
```

# Trade effects from NAFTA's tariff reductions (Table 5)

```{r}
trade <- results$trade
trade <- trade[importer %in% nafta & exporter %in% nafta]
trade <- trade[, .(trade_bln = sum(trade_bln), trade_cfl = sum(trade_cfl)),
  by = .(importer, exporter)
]
trade[, change := (trade_cfl / trade_bln - 1) * 100]
trade[exporter == importer, change := NA_real_]

trade <- dcast(
  trade,
  importer ~ exporter,
  value.var = "change"
)

trade <- trade[match(nafta, importer), ]
setcolorder(trade, c("importer", "Mexico", "Canada", "USA"))

knitr::kable(
  trade,
  digits = 2,
  latex = TRUE,
  col.names = c("Importer/Exporter", nafta),
  caption = "Trade effects from NAFTA's tariff reductions (Table 5)"
)
```

# References
